---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mregions2

<!-- badges: start -->
<!-- badges: end -->

mregions2 supersedes [mregions](https://github.com/ropensci/mregions).
It aims to:

* Retrieve marine geospatial information from the [Marineregions gazetteer](https://marineregions.org/gazetteer.php?p=webservices). 
* Access and handle data products from the [Flanders Marine Institute (VLIZ) geoserver](http://geo.vliz.be/geoserver/web/wicket/bookmarkable/org.geoserver.web.demo.MapPreviewPage?1).


## Installation

You can install the development version of mregions2 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("lifewatch/mregions2")
```

## Examples

```{r load libraries}
library(mregions2)

# helper library
library(mapview)
```

For more information about a function, type `?<function_name>` in the R console. This opens the function's documentation.

### `mr_gaz_records_by_names()`
This function retrieves records from the [Marine Gazetteer](https://marineregions.org/gazetteer.php?p=search) by inputting one or more search terms, i.e. `"North Sea"` or `"Belgium"`. The result is a [`tibble`](https://tibble.tidyverse.org) with all records that contain one of the search terms in their `preferredGazetteerName`.

```{r example mr_gaz_records_by_names()}
north_sea <- mr_gaz_records_by_names("North Sea")

head(north_sea$preferredGazetteerName)
```

<!-- When looking for the Belgian Part of the North Sea, for example, searching for `c("North Sea","Belgian")` will result in all records that contain one of the two terms. -->

<!-- ```{r example2 mr_gaz_records_by_names()} -->
<!-- northsea_belgian <- mr_gaz_records_by_names(c("North Sea", "Belgian")) -->

<!-- # Inspect all record names that were found -->
<!-- northsea_belgian$preferredGazetteerName -->
<!-- ``` -->

Records can have the same name, however, the are always distinguishable by their [MRGID](https://marineregions.org/mrgid.php). Let's look a bit closer at the records named `North Sea`.

```{r example3 mr_gaz_records_by_names()}
# north_sea_name <- north_sea[north_sea$preferredGazetteerName == "North Sea", ]
north_sea_name <- subset(north_sea, preferredGazetteerName == "North Sea")
# View(north_sea_name)
```
The 6 records with the same `preferredGazetteerName == "North Sea"` differ in their ` placeType`. The correct record to choose depends on the goal of the analysis. The status of the record can, however, help decide which one to choose. For example, the `status` of the 6 records in `north_sea_name` is either `standard`, `synonym` or `deleted`. In this case, `North Sea` with ` placeType == "IHO Sea Area"`, that has the `standard` status, should be preferred.
<!-- provide link to further info about the status names -->

```{r example4 mr_gaz_records_by_names()}
north_sea_name$placeType
north_sea_name$status
```

### `mr_gaz_record()`

This function retrieves a geometry from the [VLIZ geoserver](http://geo.vliz.be/geoserver/web/wicket/bookmarkable/org.geoserver.web.demo.MapPreviewPage?1). The geometry is a geospatial object of `class(sf)` (simple feature) from the `sf` R package. Here you can find an [introduction](https://r-spatial.github.io/sf/) to `simple feature objects` in `sf` and an [overview](https://r-spatial.github.io/sf/reference/index.html) of all available functions.

In the following example, the geometry of the North Sea IHO Sea Area is retrieved and visualised.
```{r example mr_gaz_record()}
north_sea_IHO <- subset(north_sea, placeType == "IHO Sea Area" & preferredGazetteerName == "North Sea")
mrgid <- north_sea_IHO$MRGID

# check the outcome when a geom is there and when not
# IHO_geom <- mr_gaz_record(mrgid)
# colnames(IHO_geom)
# 
# mapview(IHO_geom)
```
### `mr_gaz_info()`

This function lists information about the [Marine Gazetteer](https://marineregions.org/gazetteer.php?p=search) records, namely `placeTypes` and `sources`. This is especially useful when using `mr_gaz_records_by_type()` and `mr_gaz_records_by_source()`.The result is a [`tibble`](https://tibble.tidyverse.org) with the first 100 of all available `sources` when using `mr_gaz_info("sources")` or with all available `placeTypes` when using `mr_gaz_info("placetypes")`.

```{r example mr_gaz_info()}
all_sources <- mr_gaz_info("sources")
head(all_sources$source)
all_types <- mr_gaz_info("placetypes")
head(all_types$type)
```

### `mr_gaz_records_by_type()`

This function retrieves the first 100 records from the [Marine Gazetteer](https://marineregions.org/gazetteer.php?p=search) by inputting a placetype. This can be `"Sandbank"` or `"Archipelago"`, for instance. 
The result of `mr_gaz_records_by_type()` is a [`tibble`](https://tibble.tidyverse.org) with the first 100 records of the specified `placeType`.

In the following example, a list of all the available `placeTypes` will first be fetched using `mr_gaz_info("placetypes")`.
```{r mr_gaz_records_by_type()}
all_types <- mr_gaz_info("placetypes")

archipelagos <- mr_gaz_records_by_type("Archipelago")
head(archipelagos$preferredGazetteerName)
```

### `mr_gaz_records_by_source()`

This function retrieves the first 100 records from the [Marine Gazetteer](https://marineregions.org/gazetteer.php?p=search) by inputting a source. This can be `"GeoNames"` or `"AlgaeNames"`, for instance. In the following example, a list of the first 100 available `sources` will be fetched using `mr_gaz_info("sources")`.
The result of `mr_gaz_records_by_source()` is a [`tibble`](https://tibble.tidyverse.org) with the first 100 records of the specified `source`.

```{r mr_gaz_records_by_source()}
placetypes <- mr_gaz_info("placetypes")

archipelagos <- mr_gaz_records_by_type("Archipelago")
head(archipelagos$preferredGazetteerName)
```

### `mr_gaz_records_by_latlon()`

This function retrieves the first 100 records from the [Marine Gazetteer](https://marineregions.org/gazetteer.php?p=search) by inputting a Latitude and a Longitude.
The result of `mr_gaz_records_by_latlon()` is a [`tibble`](https://tibble.tidyverse.org) with the first 100 records of the specified `lat` and `lon`, where their centroid is within the bounding box calculated by latitude (+/- radius) and longitude (+/- radius).
A latitude and longitude radius can be given as a function argument. Check `?mr_gaz_records_by_latlon` for details.

In the following example, all records in the area of the Mariana Trench will be retrieved.

```{r example mr_gaz_records_by_latlon()}
mariana_trench <- mr_gaz_records_by_names("Mariana Trench")

mariana_trench_lat <- mariana_trench$latitude
mariana_trench_lon <- mariana_trench$longitude

records_mariana_trench <- mr_gaz_records_by_latlon(mariana_trench_lat, mariana_trench_lon)

head(records_mariana_trench$preferredGazetteerName)
```

There are many *Guyots* around Mariana Trench, apparently! Let's explore them a bit.
```{r example Guyots}
all_types <- mr_gaz_info("placetypes")

guyot <- subset(all_types, type == "Guyot")
guyot$description

guyots_mariana_trench <- subset(records_mariana_trench, placeType == "Guyot")

nrow(guyots_mariana_trench)
```
