---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mregions2

<!-- badges: start -->
<!-- badges: end -->

mregions2 supersedes [mregions](https://github.com/ropensci/mregions).
It aims to:

* Retrieve marine geospatial information from the [Marineregions gazetteer](https://marineregions.org/gazetteer.php?p=webservices). 
* Access and handle data products from the [Flanders Marine Institute (VLIZ) geoserver](http://geo.vliz.be/geoserver/web/wicket/bookmarkable/org.geoserver.web.demo.MapPreviewPage?1).


## Installation

You can install the development version of mregions2 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("lifewatch/mregions2")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(mregions2)
## basic example code
```

<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>. -->

<!-- You can also embed plots, for example: -->

<!-- ```{r pressure, echo = FALSE} -->
<!-- plot(pressure) -->
<!-- ``` -->

<!-- In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->
