---
title: "mregions2 as RDF"
author: Salvador Fernández
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mregions2-rdf}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)

options(rmarkdown.html_vignette.check_title = FALSE)
options(rdf_print_format = "turtle")
```

This vignette describes how to access the Marine Regions Gazetteer entries and geometries as RDF, which are loaded into R via `rdflib::rdf`. Mirroring the Marine Regions Gazetteer via [Linked Data Event Streams](https://w3id.org/ldes/specification) is discussed.

The following libraries are required

```{r setup, results='hide', message=FALSE}
library(mregions2)
library(rdflib)

# Use the pipe operator `%>%`
library(magrittr)

# For visualizing
library(mapview)
library(jsonlite)
```


## Retrieving the Marine Regions Gazetteer as RDF

Known an `MRGID`, you can get the gazetter entry as RDF. 

E.g. the Belgian Exclusive Economic Zone or Belgian Part of the North Sea has the `MRGID`: <https://marineregions.org/mrgid/3293> 

```{r}
gaz_search(3293) %>% 
  gaz_geometry() %>%
  mapview()
```

Using `gaz_search()` or the underlying service `gaz_rest_record_by_mrgid()` with the argument `rdf = TRUE` will:

1. Use content negotiation to get the MRGID in Turtle format. 
2. Read this document as RDF using the `rdflib` R package (`?rdflib::rdflib`) 

HTTP requests are handled with `httr2` (`?httr2::httr2`)

```{r}
bpns <- gaz_search(3293, rdf = TRUE)

class(bpns)

bpns
```

There are many stores returned. This is because the returned object includes not only info about the Gazetteer record, but also about its hierarchy. The Marine Regions Gazetteer is hierarchical and there are many relations between their records

The relations are explained the Marine Regions Ontology:

> `http://marineregions.org/ns/ontology`

There is also information about the Place Type. The Place Types are defined in:

> `http://marineregions.org/ns/placetypes`

Both the ontology and the placetypes are available as HTML when using a web browser, but also as Turtle or JSON-LD by using content negotiation.

You can then use `rdflib::rdflib` to apply common RDF applications. 

### Serialize

You can serialize the record in other formats, e.g. JSON-LD
```{r, results='hide'}
# Create a placeholder file
bpns_as_jsonld <- tempfile(fileext = ".rdf")

# Serialize
bpns %>% rdf_serialize(
  doc = bpns_as_jsonld, 
  format = "jsonld"
)

readLines(bpns_as_jsonld) %>% jsonlite::prettify()
```

<!-- Low tech solution to hide long output with HTML -->
<details><summary>Show Output</summary>
```{r, echo=FALSE}
readLines(bpns_as_jsonld) %>% jsonlite::prettify() %>% cat()
```
</details>
<!-- end -->

### SPARQL Queries

[SPARQL](https://www.w3.org/TR/sparql11-query/) queries can be applied to the RDF document. Note: The `MRGID` of the Belgian Part of the North Sea is `<http://marineregions.org/mrgid/3293>`

#### Extract all info from the record

```{r}
sparql <- "
  SELECT ?p ?o
  WHERE {
    <http://marineregions.org/mrgid/3293> ?p ?o
  }"

rdf_query(bpns, query = sparql)
```

#### Extract all relationship

The following example extracts all the predicates with the prefix `mr:`, pointing to the Marine Regions Ontology at <http://marineregions.org/ns/ontology>.

```{r}
sparql <- "
  PREFIX mr: <http://marineregions.org/ns/ontology#> 
  SELECT ?s ?p ?o
  WHERE {
    ?s ?p ?o .
    FILTER( STRSTARTS(STR(?p), str(mr:)) )
  }"

rdf_query(bpns, query = sparql)
```

#### Extract geometries

You can see in the predicates that one of the relationships is `mr:hasGeometry`, pointing to the object <http://marineregions.org/mrgid/3293/geometries?source=79&attributeValue=3293>

This URI can be requested as Turtle with content negotiation and load as an `rdflib::rdf` object of the `rdflib::rdflib` package

But you could also use `gaz_geometry()`, which already takes care of this:

```{r }
bpns_geom <- gaz_geometry(3293, format = "rdf")
bpns_geom
```

The geometry is provided as [WKT](https://en.wikipedia.org/wiki/Well-known_text_representation_of_geometry). This is the same logic that `gaz_geometry(format = "wkt")` uses to extract the geometry:

```{r}
sparql <- "
  PREFIX gsp: <http://www.opengis.net/ont/geosparql#>
  SELECT ?o
  WHERE {
    ?s gsp:asWKT ?o .
  }"

rdf_query(bpns_geom, query = sparql)
```

#### Extract the label in another language

Using the SPARQL syntax you can extract the label of the record in a specific language, if this is available. If not, the preferred label will be returned

Example: Get the Gazetteer entry for _"Belgium"_ with MRGID `<http://marineregions.org/mrgid/14>` and extract the label in Dutch (nl) if available.

```{r}
belgium <- gaz_search(14, rdf = TRUE)

sparql <- '
  PREFIX skos: <http://www.w3.org/2004/02/skos/core#>
  SELECT (COALESCE(?altLabel, ?prefLabel) AS ?o)
  WHERE {
    OPTIONAL {
      ?s skos:altLabel ?altLabel .
      filter langMatches(lang(?altLabel), "nl")
    }
    ?s skos:prefLabel ?prefLabel .
  }
'

rdf_query(belgium, query = sparql)
```

## Mirror the Marine Regions Gazetteer using the LDES Feed

The Marine Regions Gazetteer is accessible as RDF, and can be mirrored and synchronized via [Linked Data Event Streams (LDES)](https://w3id.org/ldes/specification). This is described with high detail in: 

> Lonneville B. et al. (2021) Publishing the Marine Regions Gazetteer as a Linked Data Event Stream. S4BioDiv 2021. http://ceur-ws.org/Vol-2969/paper8-s4biodiv.pdf

Building such API in R is, at the moment, out of the scope of mregions2. However, an interested user could make use of mregions2 and `rdflib::rdflib` to mirror the Marine Regions Gazetteer. The general approach is described below:

### Initialize a database

By design, mregions2 stores the RDF triples in memory. This is because the package does not aim at replicating the Marine Regions Gazetteer, but only consult small subsets. 

`rdflib::rdflib` allows to store the triples in a database. This is further explained in `vignette("storage", package = "rdflib")`.

As an example, using a [Virtuoso](https://en.wikipedia.org/wiki/Virtuoso_Universal_Server) database backend is possible. Virtuoso is a popular open source database for RDF.

```r
triplestore <- rdf(storage = "virtuoso", 
                   user = "dba", 
                   password = "dba", 
                   dsn = "Local Virtuoso"
                   )

# Turn bpns into nquads
tmp <- tempfile(fileext = ".nq")
write_nquads(bpns, tmp)

# Parse bpns as nquads into the store
rdf_parse(tmp, format = "nquads", rdf = triplestore)
```

### Linked Data Event Stream Feed

The LDES feed is available at:

> `https://marineregions.org/feed`

This offers fragmentation of the whole Marine Regions Gazetteer based on the last modification time of the `?MRGID`. It primary use is to make it easy to have external caches, replications, or derived indexes in an efficient and incremental way.

The implementation would start by fetching the feed as start page and then follow the `tree:relation` paths down to pages that contain only `tree:member` entries that are older than the requested date.

<sup>Source: https://github.com/ropensci/mregions/pull/62#issuecomment-1091837752</sup>


```{r rdf, include = FALSE, echo=FALSE}
library(mregions2)

#
#
#
#
#
#
#
#
#
#
# cuba_eez_relations <- mr_relations(cuba_eez)
#
# cuba_eez %>% tidyr::as_tibble() %>% View()
#
# cuba_eez <- mr_get(8406)
#
#
# test <- rdflib::rdf_query(cuba_eez, query = paste0(
#   mr_prefixes(),
#   "
#   SELECT *
#   WHERE {
#     <http://marineregions.org/mrgid/8406> ?p ?o
#
#   }
#   "
#
# ))
#
#
# ## SHOW TO THE TEAM DIFFERENT OPTIONS AND DECIDE WHAT TO DO
# ## OPTION 1: STICK TO REST WEBSERVICES AND PROVIDE LINKED DATA AS VIGNETTE
# ## OPTION 2: PROVIDE ALL STUFF IN A GEOOBJECT AS A TABLE
# ## OPTION 3: CURATE ALL STUFF IN A GEOOBJECT AS TABLES
#
#
# test = cbind(s = "http://marineregions.org/mrgid/8406", test)
#
# test = rbind(test, c("http://marineregions.org/mrgid/845456",
#                      "foo_name",
#                     "foo value"))
#
# colnames(test)
#
# tidyr::pivot_wider(test, names_from = p, values_from = o) %>% View()
#
#
#
#
#
#
# cuba_eez %>%
#   rdflib::rdf_serialize(format = "turtle", namespace = c(
#     mr = "http://marineregions.org/ns/ontology#",
#     mrt = "http://marineregions.org/ns/placetypes#",
#     dc = "http://purl.org/dc/terms/",
#     xsd = "http://www.w3.org/2001/XMLSchema#",
#     rdfs = "http://www.w3.org/2000/01/rdf-schema#",
#     skos = "http://www.w3.org/2004/02/skos/core#",
#     dcat = "http://www.w3.org/ns/dcat#",
#     gsp = "http://www.opengis.net/ont/geosparql#",
#     prov = "http://www.w3.org/ns/prov#"
#   )) %>%
#   cat()
# # as_tibble.MRGeoObject(cuba_eez) %>% View()
#
#
#
# View(cuba_eez_relations)
#
#
# # Do something with replaced by
# bouvet <- mr_get(8399) %>% as.data.frame()
#
#
#
#
# mr_get("http://marineregions.org/mrgid/8406") %>%
#   rdflib::rdf_serialize(format = "turtle", namespace = c(
#     mr = "http://marineregions.org/ns/ontology#",
#     mrt = "http://marineregions.org/ns/placetypes#",
#     dc = "http://purl.org/dc/terms/",
#     xsd = "http://www.w3.org/2001/XMLSchema#",
#     rdfs = "http://www.w3.org/2000/01/rdf-schema#",
#     skos = "http://www.w3.org/2004/02/skos/core#",
#     dcat = "http://www.w3.org/ns/dcat#",
#     gsp = "http://www.opengis.net/ont/geosparql#",
#     prov = "http://www.w3.org/ns/prov#"
#   )) %>%
#   cat()
#
#
#
# belgium <- mr_get("http://marineregions.org/mrgid/26567")
#
#
#
#
#
#
#
#
#
#
# query <- paste0(
#   "
#   prefix mr: <http://marineregions.org/ns/ontology#>
#   prefix mrt: <http://marineregions.org/ns/placetypes#>
#   prefix dc: <http://purl.org/dc/terms/>
#   prefix xsd: <http://www.w3.org/2001/XMLSchema#>
#   prefix rdfs: <http://www.w3.org/2000/01/rdf-schema#>
#   prefix skos: <http://www.w3.org/2004/02/skos/core#>
#   prefix dcat: <http://www.w3.org/ns/dcat#>
#   prefix gsp: <http://www.opengis.net/ont/geosparql#>
#   prefix prov: <http://www.w3.org/ns/prov#>
#
#   SELECT ?subject ?predicate ?relation
#
#          # ?inContextOf ?isRelatedTo ?isPartOf
#          # ?contains ?isAdministrativePartOf ?hasAdministrativePart
#          # ?isPartlyPartOf ?partlyContains ?flowsInto ?hasTributary
#          # ?risesIn ?containsSourceOf ?flowsThrough ?containsWatercourse
#          # ?isInfluencedBy ?influences
#
#
#   WHERE {
#
#     # Get label
#     ?mrgid a ?placetype ;
#            skos:prefLabel ?label ;
#            dc:modified ?modified .
#
#     # Restrict to only the placetype
#     ?placetype skos:inScheme <http://marineregions.org/ns/placetypes> .
#
#     # modified is the filter - it is or should be in all mrgid
#
#     # OPTIONAL { ?mrgid mr:inContextOf ?inContextOf }
#     # OPTIONAL { ?mrgid mr:isRelatedTo ?isRelatedTo }
#     # OPTIONAL { ?mrgid mr:isPartOf ?isPartOf }
#     # OPTIONAL { ?mrgid mr:contains ?contains }
#     # OPTIONAL { ?mrgid mr:isAdministrativePartOf ?isAdministrativePartOf }
#     # OPTIONAL { ?mrgid mr:hasAdministrativePart ?hasAdministrativePart }
#     # OPTIONAL { ?mrgid mr:isPartlyPartOf ?isPartlyPartOf }
#     # OPTIONAL { ?mrgid mr:partlyContains ?partlyContains }
#     # OPTIONAL { ?mrgid mr:flowsInto ?flowsInto }
#     # OPTIONAL { ?mrgid mr:hasTributary ?hasTributary }
#     # OPTIONAL { ?mrgid mr:risesIn ?risesIn }
#     # OPTIONAL { ?mrgid mr:containsSourceOf ?containsSourceOf }
#     # OPTIONAL { ?mrgid mr:flowsThrough ?flowsThrough }
#     # OPTIONAL { ?mrgid mr:containsWatercourse ?containsWatercourse }
#     # OPTIONAL { ?mrgid mr:isInfluencedBy ?isInfluencedBy }
#     # OPTIONAL { ?mrgid mr:influences ?influences }
#
#     OPTIONAL {
#     { <http://marineregions.org/mrgid/8406> mr:contains  ?relation } UNION { <http://marineregions.org/mrgid/8406> mr:partlyContains ?relation  }
#     }
#
#     OPTIONAL {
#     { ?mrgid mr:contains  ?relation } UNION { ?mrgid mr:partlyContains ?relation  }
#     }
#
#   }"
# )
#
#
#
#
# query <- paste0(
#   "
#   prefix mr: <http://marineregions.org/ns/ontology#>
#   prefix mrt: <http://marineregions.org/ns/placetypes#>
#   prefix dc: <http://purl.org/dc/terms/>
#   prefix xsd: <http://www.w3.org/2001/XMLSchema#>
#   prefix rdfs: <http://www.w3.org/2000/01/rdf-schema#>
#   prefix skos: <http://www.w3.org/2004/02/skos/core#>
#   prefix dcat: <http://www.w3.org/ns/dcat#>
#   prefix gsp: <http://www.opengis.net/ont/geosparql#>
#   prefix prov: <http://www.w3.org/ns/prov#>
#
#   SELECT ?mrgid
#          (COALESCE(?altLabel, ?prefLabel) AS ?label)
#          ?placetype
#          ?modified
#          ?source
#          ?seeAlso
#          ?centroid
#          ?bbox
#          ?the_geom
#
#   WHERE {
#
#     # Restrict to only the placetype
#     ?mrgid a ?placetype .
#     ?placetype skos:inScheme <http://marineregions.org/ns/placetypes> .
#
#     # Get some vars
#     OPTIONAL { ?mrgid dc:modified ?modified }
#     OPTIONAL { ?mrgid rdfs:seeAlso ?seeAlso }
#     OPTIONAL { ?mrgid dcat:centroid ?centroid }
#     OPTIONAL { ?mrgid dcat:bbox ?bbox }
#     OPTIONAL { ?mrgid mr:hasGeometry ?the_geom }
#
#
#     # Get the language
#     OPTIONAL {
#       ?mrgid skos:altLabel ?altLabel .
#       filter langMatches(lang(?altLabel), \"nl\")
#     }
#     ?mrgid skos:prefLabel ?prefLabel .
#
#     # Get the source
#     OPTIONAL { ?mrgid prov:hadPrimarySource ?primarySource }
#     OPTIONAL { ?primarySource prov:wasAttributedTo ?wasAttributedTo }
#     OPTIONAL { ?wasAttributedTo rdfs:label ?source }
#
#     # Approach using Property Paths, but seem like they are not supported
#     # ?mrgid prov:hadPrimarySource/prov:wasAttributedTo/rdfs:label ?source .
#
#
#
#   }"
# )
#
# x <- rdflib::rdf_query(belgium, query)
#
#
# # Add base_graph
#
#
#
#
#
#
#
#
#
# # debug(mr_get_rdf)
# options(verbose = F)
#
# belgium <- mr_get("http://marineregions.org/mrgid/14")
# tidyr::as_tibble
#
#
#
#
#
# france <- mr_get("http://marineregions.org/mrgid/17")
#
# mr_
#
#
# ## CREATE GENERAL FUNCTION TO MERGE MRGIDS
# ## CREATE GENERAL FUNCTION TO APPLY SPARQL QUERY
#
# ## QUERIES
# ## GET ALL BASIC INFO
#
#
#
#
#
# print(out, n = 100)
#
# #
# # "
# # prov:hadPrimarySource [
# #   prov:wasAttributedTo [
# #     rdfs:label
# #       \"(2001). The Times comprehensive atlas of the world. 10th ed. Times Books:
# #       London. ISBN 0-7230-0792-6. 67, 220, 124 plates pp.\"^^xsd:string
# #   ]
# # ]
# #
# # "
#
#
#
#
#
#
#
# ## GET ALL RELATIONSHIPS
# query <- paste0(
#   prefix,
#   "
#   SELECT ?mrgid ?placetype ?prefLabel
#   WHERE {
#     ?mrgid a ?placetype ;
#           skos:prefLabel ?prefLabel .
#
#   }"
# )
#
# cat(query)
# out <- rdflib::rdf_query(rdf, query, data.frame = FALSE)
# out <- subset(out, out$placetype != 'http://marineregions.org/ns/ontology#MRGeoObject')
# out <- subset(out, out$mrgid != 'http://marineregions.org/mrgid/14')
# print(out, n = 100)
#
# query <- paste0(
#   prefix,
#   "
#   SELECT ?relationship ?value
#   WHERE {
#     <http://marineregions.org/mrgid/14> ?relationship ?value .
#   }"
# )
#
# relations <- rdflib::rdf_query(rdf, query, data.frame = FALSE)
# relations <- subset(relations, grepl("http://marineregions.org/ns/ontology#", relations$relationship, fixed = TRUE))
#
# out <- subset(out, out$placetype != 'http://marineregions.org/ns/ontology#MRGeoObject')
# out <- subset(out, out$mrgid != 'http://marineregions.org/mrgid/14')
# print(out, n = 100)
#
#
#
#
# ## GET GEOMETRY
#
#
#
#
#
#
#
#
# ## GET PREFERRED LANGUAGE
# query <- paste0(
#   prefix,
#   "
#   SELECT (<http://marineregions.org/mrgid/14> as ?mrgid)
#          (coalesce(?altLabel, ?prefLabel) as ?label)
#   WHERE {
#
#
#     OPTIONAL {
#       <http://marineregions.org/mrgid/14> skos:altLabel ?altLabel
#       filter langMatches(lang(?altLabel), \"nl\")
#     }
#     OPTIONAL {
#       <http://marineregions.org/mrgid/14> skos:prefLabel ?prefLabel
#     }
#
#   }
#
#
#   "
# )
#
# cat(query)
# out <- rdflib::rdf_query(rdf, query, data.frame = FALSE)
# out
#
#
#
# #
# # cat(cont)
# # query <- paste0(
# #   prefix,
# #   "
# #   SELECT ?mrgid ?name ?placetype
# #   WHERE {
# #     ?mrgid a ?placetype .
# #     ?mrgid skos:prefLabel ?name .
# #   }"
# # )
# #
# # cat(query)
# # rdflib::rdf_query(rdf, query, data.frame = FALSE)
#
#
# cat(cont)
# query <- paste0(
#   prefix,
#   "
#   SELECT ?mrgid ?label ?geoobject
#   WHERE {
#     ?mrgid a ?geoobject .
#     ?mrgid skos:altLabel ?label .
#     FILTER (lang(?label) = 'en')
#   }
#   LIMIT 15
#   "
# )
#
# cat(query)
# rdflib::rdf_query(rdf, query, data.frame = FALSE)
#
#
# # <http:\/\/marineregions.org\/ns\/ontology#(\w+)>
# # FILTER (regex(?predicate, '/<http:\\/\\/marineregions.org\\/ns\\/ontology#(\\w+)>/g')) .
#
# FILTER (regex(?predicate, '/mr:(\\w+)/g')) .
#
#
# # Set prefixes
# todelete <- gsub("@prefix(.*?)\n", "", resp)
# prefix <- gsub(todelete, "", resp, fixed = TRUE)
# prefix <- gsub("@", "", prefix, fixed = TRUE)
# prefix <- gsub(" .", "", prefix, fixed = TRUE)
# Sys.setenv(mr_prefix = prefix)
#
#
#
# query <-
#   "
#   PREFIX mr: <http://marineregions.org/ns/ontology#>
#   SELECT ?mrgid ?the_geom
#   WHERE { ?mrgid mr:hasGeometry ?the_geom .}"
#
#
# uri_geom <- rdflib::rdf_query(rdf, query)
# #
# # query <-
# #   "
# #   PREFIX mr: <http://marineregions.org/ns/ontology#>
# #   SELECT ?a ?c
# #   WHERE { ?a mr:contains ?c .}"
# #
# # contains <- rdflib::rdf_query(rdf, query, data.frame = FALSE)
#
#
# query <-
#   "
#   PREFIX mr: <http://marineregions.org/ns/ontology#>
#   PREFIX skos: <http://www.w3.org/2004/02/skos/core#>
#   SELECT ?mrgid ?contains ?name ?other_names ?the_geom
#   WHERE { ?mrgid mr:contains ?contains .
#           ?mrgid skos:prefLabel ?name .
#           ?mrgid skos:altLabel ?other_names .
#           ?mrgid mr:hasGeometry ?the_geom .
#         }
# "
#
# contains <- rdflib::rdf_query(rdf, query, data.frame = T)
#
#
# # How to select the skos:prefLabel from the relationships
# # How to get the placetype?
# # Create method to jump to next?
#
#
# "http://marineregions.org/ns/ontology" # no cacheable
# "http://marineregions.org/ns/placetypes.ttl" # no cacheable
# "https://www.w3.org/ns/dcat2" # cache-control + expires
# "http://www.w3.org/ns/prov" # cache-control + expires
# "http://www.w3.org/2000/01/rdf-schema" # cache-control + expires
# "http://www.opengis.net/ont/geosparql" # Last-Modified + ETag
#
# 'http://purl.org/dc/terms/'
# "http://purl.org/dc/elements/1.1/"
#
#' #' Create a new S3 object of class "MRGeoObject"
#' #'
#' #' Inherits the class "rdf" from the package "rdflib"
#' #'
#' #' @param x
#' #'
#' #' @return
#' #'
#' #' @examples
#' new_MRGeoObject <- function(x = list()){
#' 
#'   structure(
#'     x,
#'     class = c("MRGeoObject", "rdf")
#'   )
#' 
#' }
#' 
#' #' Get a Marine Regions GeoObject as RDF
#' #'
#' #' @param x a url, a mrgid or a name of a Marine Regions Geoobject as a string
#' #' @param ... aditional params for furter inheritance
#' #'
#' #' @return an object of class c("MRGeoObject", "rdf")
#' #' This is completely compatible with the rdflib package.
#' #' For instance, you can apply sparql queries with rdflib::rdf_query()
#' #'
#' #' Or you can serialize the file with rdflib::rdf_serialize()
#' #'
#' #' Add extra RDF documents to and create your own knowledge graph
#' #'
#' #' @export
#' #'
#' #' @examples
#' #' x <- mr_get("Belgium")
#' #' x <- mr_get("http://marineregions.org/mrgid/14)
#' #' x <- mr_get(14)
#' mr_get <- function(x, ...){
#'   UseMethod("mr_get")
#' }
#' 
#' #' @rdname mr_get
#' mr_get.default <- function(url, ...){
#' 
#'   resp <- mr_request(url) %>%
#'     httr2::req_headers(accept = "text/turtle") %>%
#'     httr2::req_perform("GET") %>%
#'     httr2::resp_body_string()
#' 
#'   verbose <- getOption("verbose", default = FALSE)
#'   if(verbose){
#'     cat(resp)
#'   }
#' 
#'   out <- resp %>%
#'     rdflib::rdf_parse("turtle")
#' 
#'   new_MRGeoObject(out)
#' 
#' }
#' 
#' #' @rdname mr_get
#' mr_get.numeric <- function(mrgid, ...){
#'   url <- glue::glue("http://marineregions.org/mrgid/{mrgid}")
#'   mr_get.default(url)
#' }
#' 
#' #' @rdname mr_get
#' mr_get.character <- function(string, ...){
#' 
#'   is_mr_url <- grepl("http://marineregions.org", string, fixed = TRUE)
#' 
#'   if(is_mr_url){
#'     out <- mr_get.default(string)
#'     return(out)
#'   }
#' 
#'   if(!is_mr_url){
#' 
#'     search <- mregions2::mr_gaz_records_by_names(string)
#' 
#'     if(nrow(search) == 0){
#'       warning(glue::glue("No matches found for \"{string}\""), .call = FALSE)
#'       return(invisible(NULL))
#'     }
#' 
#'     verbose <- getOption("verbose", default = FALSE)
#'     if(verbose){
#'       cli::cli_alert_warning(glue::glue("{nrow(search)} match(es) found for \"{string}\": '{paste0(search$preferredGazetteerName, collapse = '\', \'')}'"))
#'     }
#' 
#'     return(mr_get.numeric(search$MRGID))
#'   }
#' }
#' 
#' 
#' #' Returns all the prefixes for sparql queries
#' #'
#' #' @examples
#' #' mr_prefixes()
#' mr_prefixes <- function(){
#'   prefixes <- c(
#'   "prefix mr: <http://marineregions.org/ns/ontology#>",
#'   "prefix mrt: <http://marineregions.org/ns/placetypes#>",
#'   "prefix dc: <http://purl.org/dc/terms/>",
#'   "prefix xsd: <http://www.w3.org/2001/XMLSchema#>",
#'   "prefix rdfs: <http://www.w3.org/2000/01/rdf-schema#>",
#'   "prefix skos: <http://www.w3.org/2004/02/skos/core#>",
#'   "prefix dcat: <http://www.w3.org/ns/dcat#>",
#'   "prefix gsp: <http://www.opengis.net/ont/geosparql#>",
#'   "prefix prov: <http://www.w3.org/ns/prov#>"
#'   )
#' 
#'   paste0(prefixes, collapse = "\n")
#' 
#' }
#' 
#' 
#' 
#' #' Extract information from a Marine Regions Geo Object as tabular data
#' #'
#' #' @param x an object of class c("MRGeoObject", "rdf")
#' #' @param lang preferred language as 2c ISO_639_1. See details.
#' #' @param ... aditional params for furter inheritance
#' #'
#' #' @details
#' #'
#' #' The Marine Regions Geo Objects can be labels in several
#' #' languages. Choose your preferred language and, if the label exists
#' #' in your language, it will be returned. Otherwise the label will be in the preferred language.
#' #'
#' #' @return a tibble with one row per Marine Regions Geo Object
#' #' @export
#' #'
#' #' @examples
#' as_tibble.MRGeoObject <- function(x, lang = "en", ...){
#' 
#'   checkmate::assert_choice(lang, c("en", "es", "fr", "de", "nl", "it"))
#' 
#'   query <- paste0(
#'     mr_prefixes(),
#'     "
#'   SELECT ?mrgid
#'          (COALESCE(?altLabel, ?prefLabel) AS ?label)
#'          ?placetype
#'          ?modified
#'          ?source
#'          ?seeAlso
#'          ?centroid
#'          ?bbox
#'          ?the_geom
#'
#'   WHERE {
#'
#'     # Restrict to only the placetype
#'     ?mrgid a ?placetype .
#'       FILTER( STRSTARTS(STR(?placetype),str(mrt:)) )
#'
#'     # modified is the filter - it is or should be in all mrgid
#'     ?mrgid dc:modified ?modified
#'     OPTIONAL { ?mrgid rdfs:seeAlso ?seeAlso }
#'     OPTIONAL { ?mrgid dcat:centroid ?centroid }
#'     OPTIONAL { ?mrgid dcat:bbox ?bbox }
#'     OPTIONAL { ?mrgid mr:hasGeometry ?the_geom }
#'
#'
#'     # Get the language
#'     OPTIONAL {
#'       ?mrgid skos:altLabel ?altLabel .
#'       filter langMatches(lang(?altLabel), \"nl\")
#'     }
#'     ?mrgid skos:prefLabel ?prefLabel .
#'
#'     # Get the source
#'     OPTIONAL { ?mrgid prov:hadPrimarySource ?primarySource }
#'     OPTIONAL { ?primarySource prov:wasAttributedTo ?wasAttributedTo }
#'     OPTIONAL { ?wasAttributedTo rdfs:label ?source }
#'
#'   }"
#'   )
#' 
#'   # Perform Query
#'   out <- rdflib::rdf_query(x, query) %>%
#'     suppressMessages()
#' 
#'   out
#' }
#' 

#' 
#' options(verbose = TRUE)
#' 
#' cuba_eez <- mr_get("Cuban Exclusive Economic Zone")
#' 
#' tidyr::as_tibble(cuba_eez) %>% View()
#' 
#' mr_relations(cuba_eez) %>% View()
#' 
#' test1 <- mregions2::mr_gaz_relations_by_MRGID(8406,
#'                                      direction = "both",
#'                                      type = "all")
#' View(test1)
#' 
#' #' @rdname as_tibble.MRGeoObject
#' as.data.frame.MRGeoObject <- function(x, ...){
#'   x <- as_tibble.MRGeoObject(x)
#'   as.data.frame(x)
#' }
#' 
#' 
#' #' Extract the relationships of a given MRGeoObject
#' #'
#' #' @param x a MRGeoObject
#' #' @param ...
#' #'
#' #' @return
#' #' @export
#' #'
#' #' @examples
#' #' cuba_eez_relationships <- mr_get(8406) %>% mr_relationships()
#' mr_relations <- function(x, ...){
#'   # 1. Get all predictes of ontology mr:
#'   # These are the relations plus hasGeometry
#'   query <- paste0(
#'     mr_prefixes(),
#'     "
#'   SELECT ?s ?relation ?mrgid
#'   WHERE {
#'     ?s ?relation ?mrgid .
#'     FILTER( STRSTARTS(STR(?relation),str(mr:)) )
#' 
#'   }"
#'   )
#' 
#'   suppressMessages(out <- rdflib::rdf_query(x, query, data.frame = FALSE))
#' 
#'   # Leave out geom, no idea how to do it in SPARQL
#'   # out <- subset(out, out$relation != "http://marineregions.org/ns/ontology#hasGeometry")
#' 
#'   # 2. Get all placetypes and labels of relations
#'   query2 <- paste0(
#'     mr_prefixes(),
#'     "
#'     SELECT ?mrgid ?placetype ?label
#'     WHERE {
#'       ?mrgid a ?placetype.
#'       FILTER( STRSTARTS(STR(?placetype),str(mrt:)) )
#' 
#'       ?mrgid skos:prefLabel ?label.
#' 
#' 
#'     }
#'     "
#'   )
#' 
#'   suppressMessages(labels <- rdflib::rdf_query(x, query2, data.frame = FALSE))
#' 
#' 
#'   # 3. Merge dataframes, rename and rearrange to provide human-readable dataset
#'   out <- merge(out, labels)
#' 
#'   out <- data.frame(
#'     relation = gsub("http://marineregions.org/ns/ontology#", "mr:",  out$relation),
#'     label = out$label,
#'     placetype = gsub("http://marineregions.org/ns/placetypes#", "mrt:", out$placetype),
#'     mrgid = out$mrgid
#'   )
#' 
#'   out <- out[order(out$relation, decreasing = FALSE), ]
#' 
#'   # End
#'   tidyr::as_tibble(out)
#' 
#' }
```
