---
title: "mregions2 package development info"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{packagedev_june2022}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This vignette serves to document the status of the `mregions2` package development process in June, 2022 after the IMBRSea professional practice internship of @whaleshark99 at VLIZ, Belgium.

## package structure

All relevant functions from the gazetteer webservices are in `mr_gaz_records.R`.
All of them contain assertions from the `checkmate`-package. Assertions are not claimed to be complete so it might be good to check that assertions are there for each step.

Tests with the `testthat`-package were started with but definetely have to be completed/altered. 

I most recently worked in the branch `dev_OOP` that I did not merge into `main`. On this branch I created the class `MarineRegion` using the `R6`-package.
On this branch, there is a function called `mr_gaz_relations_full_by_MRGID` in `mr_gaz_records.R`, which does not use the webservices but fetches the relations from the open-linked data through `mr_gaz_ldes()`. This might be useful to include in the general package. The function is included at the very end of the vignette.
 

## wrapup meeting 01.06.2022

### points of discussion

 - remove `_gaz_` in function names?
 - 

### agreed changes

#### conventional approach

 - [ ] rename `mr_gaz_ldes()` to something more informative. Potentially display information in a clearer (i.e. less nested) way.

 
#### Object Oriented Programming (OOP) approach

 - [ ] rename `mr_marine_region()` to `mr_GeoObject()` and the class `MarineRegion` to `GeoObject`.
 - [ ] remove `area` in the object creation of `mr_marine_region()`
 - [ ] change the description of a `marine_region` in the RStudio environment (now its `R6 Object containing active binding0`)

## Next Steps

 - [ ] explore and include lazy loading
 - [ ] `mr_gaz_relations_full_by_MRGID()`: change the way a relation is retrieved from `mr_gaz_ldes()`.
 - [ ] make CRS accessible in `gaz_geometry()`
 - [ ] print comprehensible error message when PC is not connected to internet. I started with this in `test-mr_gaz_records.R` but the test does not make much sense just yet.
 - [ ] include sth that checks the language of a gazetteer entry
 - [ ] make clearer function name for `mr_gaz_record()`. Here, the user fetches the geometry (if existent). Transport this information through the function name, e.g. by changing the name to `mr_gaz_record_geom()`?
 

## Things that I think could be nice in the package

- [ ] enable km radius in `getGazetteerRecordsByLatLong`. At the moment, a lat_radius and lon_radius of 0 gives 100 results (the max retrievable number) because its units are in degrees. 
- [ ] create a wrapper function of existing webservice functions that lets the user specify more attributes of a gazetteer record, e.g. `name == "France" & placeType == "Nation" & language == "English"`. Yes, this can also be done in an additional step by the user itself but could speed up the workflow of retrieving a record. 
- [ ] add column to a gazetteer record that specifies if that record has a geometry or not. `mr_has_geometry()` already exists but it might be helpful to give that information to the user directly without requiring them to know this function. In addition, it might prevent confusion when someone tries to get a geometry and an error is thrown.


```{r fun mr_gaz_relations_full_by_MRGID}
#' Get all Relations for a Marine Gazetteer Record
#'
#' @param mrgid The Marine Regions Geographic IDentifier.
#'
#' @return A `tibble` with all records that are related. The column `relation` contains the type of relation.
#' Explanations of the relation types can be found at: \url{https://marineregions.org/ontology/documentation.html#objectproperties}.
#' @export
#' @examples
#' belgium_mrgid <- 14
#' belgium_relations <- mr_gaz_relations_full_by_MRGID(belgium_mrgid)
mr_gaz_relations_full_by_MRGID <- function(mrgid){

  record <- mregions2::mr_gaz_ldes(mrgid, type = "list")

  relations = record[grepl("mr:", names(record))]
  if("mr:hasGeometry" %in% names(relations)) relations[["mr:hasGeometry"]] <- NULL

  for (i in 1:length(relations)) {
    for (j in 1:length(relations[[i]])) {
      relative_url <- relations[[i]][[j]][["@id"]]
      relative_mrgid <- gsub("http://marineregions.org/mrgid/", "", relative_url)
      relative_mrgid <- as.integer(relative_mrgid)
      temp_res_j <- mregions2::mr_gaz_record(relative_mrgid, add_geometry = FALSE)
      temp_res_j$relation <- gsub("mr:", "", names(relations[i]))

      ifelse(j == 1,
             res_j <- temp_res_j,
             res_j <- rbind(res_j, temp_res_j))
    }
    ifelse(i == 1,
           res <- res_j,
           res <- rbind(res, res_j))
  }

  return(res)

  # TODO: add assertions and unit tests
  # TODO: turn for loops into lapply functions
}

```

