---
title: "Marine Regions Data Products Ontology"
author: Salvador Fernández
output: 
  html_document:
    df_print: tibble
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
 warning = FALSE,
 message = FALSE
)
```

> An ontology is a categorization of all of the concepts in some field of knowledge.

<small>[GNU Collaborative International Dictionary of English](https://gcide.gnu.org.ua)</small>

The data products hosted at marineregions.org are of vector type, which they come together with attribute tables. These attribute tables are named, but such names are not yet standardized.

To ease the understanding of the data products, the Marine Regions Team has described the attributes.

The Flanders Marine Institute (VLIZ), hosting institution behind Marine Regions, is currently working towards Open Linked Data standards ([Lonneville et al., 2021](http://ceur-ws.org/Vol-2969/paper8-s4biodiv.pdf)). The naming of these attributes will likely turn into standardized terms reusing or expanding on-going RDF-available ontologies.

This work was triggered during the [rOpenSci review](https://github.com/ropensci/software-review/issues/590#issuecomment-1606138219). Thanks to [\@sheilasaia](https://github.com/sheilasaia) for raising this issue.

You can consult the list in your R session via the `mrp_ontology` object that is loaded with `mregions2`

```{r}
library(mregions2)
mrp_ontology
```

---

```{r results='asis', echo = FALSE}
library(dplyr)

for(i in 1:nrow(mrp_list)){
  cat(glue::glue("### {mrp_list$title[i]}"))
  cat("\n")
  cat(glue::glue("{mrp_list$abstract[i]}"))
  cat("\n")
  df <- mrp_ontology %>% filter(layer == mrp_list$layer[i])
  print(knitr::kable(df))
  cat("\n")
}


```

